### 2015-10-29 AtmoSwing - Geoscientific Model Development/template.tex
Using the AM abbreviation.

%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
%% ---------------------------------
%% This template should be used for copernicus.cls
%% The class file and some style files are bundled in the Copernicus Latex Package which can be downloaded from the different journal webpages.
%% For further assistance please contact the Copernicus Publications at: publications@copernicus.org
%% http://publications.copernicus.org

%% Please use the following documentclass and Journal Abbreviations for Discussion Papers and Final Revised Papers.

%% 2-Column Papers and Discussion Papers
%\documentclass[gmd, manuscript]{copernicus}
\documentclass[gmd]{copernicus}

%% Journal Abbreviations (Please use the same for Discussion Papers and Final Revised Papers)

% Geoscientific Model Development (gmd)
% Hydrology and Earth System Sciences (hess)

%% \usepackage commands included in the copernicus.cls:
%\usepackage[german, english]{babel}
%\usepackage{tabularx}
%\usepackage{cancel}
%\usepackage{multirow}
%\usepackage{supertabular}
%\usepackage{algorithmic}
%\usepackage{algorithm}
%\usepackage{float}
%\usepackage{subfig}
%\usepackage{rotating}

\begin{document}

\linenumbers

\title{AtmoSwing (v1.4): Analog Technique Model for Statistical weather forecastING}

% \Author[affil]{given_name}{surname}

\Author[1,2]{Pascal}{Horton}
\Author[1]{Michel}{Jaboyedoff}
\Author[3]{Charles}{Obled}

\affil[1]{University of Lausanne, Lausanne, Switzerland}
\affil[2]{Terranum, Lausanne, Switzerland}
\affil[3]{Universit\'{e} de Grenoble-Alpes, LTHE, Grenoble, France}

%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.

\runningtitle{AtmoSwing}

\runningauthor{P. Horton et al.}

\correspondence{Pascal Horton (pascal.horton@alumnil.unil.ch)}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

%% These dates will be inserted by Copernicus Publications during the typesetting process.

\firstpage{1}

\maketitle

\begin{abstract}
% TODO: write

\end{abstract}

\introduction  %% \introduction[modified heading if necessary]
% TODO: write

\section{The analogue method}

As AtmoSwing does not rely heavily on one variant of the analogue method (AM), but can implement different parametrization, we gathered here several references of possible implementations as well as the main evolution steps of the AM. This latter summarizes the improvements that were brought, as well as discontinued approaches that were found less relevant (such as the use of PCA on the predictors).

\subsection{Principles}

The AM is based on the principle that two similar synoptic situations may produce similar local effects \citep{Lorenz1956}. The perfect analogy does not exist, but sufficiently similar situations leading to similar effects can be identified. Thus, two states of the atmosphere that are alike are called analogues \citep{Lorenz1969}. To be relevant, this analogy must be selected by optimizing the following elements:

\begin{itemize}
	\item The analogy variable (the predictor) must contain synoptic information having indirect dependency with the local weather and the target variable (often the precipitation).
	\item The spatial window is the domain on which the analogy variable is considered. The ideal size of this area is one that maximizes useful information.
	\item The temporal window is the hour of day at which we consider the analogy variable. The analogy can be instantaneous (eg. 12~h~UTC) or averaged over a period (eg. 24~h).
	\item The analogy period is the time of year in which similar situations are being sought. Thus, we compare situations with a similar distribution of solar energy \citep{Lorenz1969}.
	\item The analogy criterion, needed to compare the variables on the chosen spatial and temporal windows, is a score used for ranking past situations according to their degree of similarity with the target situation \citep{Bontron2004}.
	\item The ideal number of analogues situations is the best compromise in order to take into account local variability and to maximize useful synoptic information  \citep{Bontron2004}.
\end{itemize}

Because of the chaotic nature of the atmosphere, two analogue situations quickly diverge over time \citep{Lorenz1969}. Thus, the AM has strong limitations regarding the temporal extrapolation \citep{Bontron2004}. Numerical models being more capable of simulating the dynamic evolution of the atmosphere, the temporal extrapolation of the synoptic variables is left to them. The search for analogy aims thus at connecting the forecasted synoptic situation with a local predictand (temperature, precipitation ...), which is more difficult to simulate by numerical models.

The AM is being used for many years at EDF (\'{E}lectricit\'{e} de France), to build precipitation and temperature scenarios \citep{Desaint2008a}. It is also operational at the Compagnie Nationale du Rh\^{o}ne (CNR, France) and some flood forecasting services (SPC) in France and Switzerland \citep{Marty2010,GarciaHernandez2009b,Horton2012}.

\subsection{Data}
\label{sec:data}

The AM relies on two types of data: predictors, that are atmospheric variables describing the state of the atmosphere at a synoptic scale, and the predictand, which is the local weather time series we want to forecast.

Predictors are generally reanalysis datasets (as archive) and outputs of a global numerical weather prediction models for the target situation in operational forecasting. Reanalysis datasets, such as the NCEP/NCAR reanalysis I \citep[6-hourly, 17 pressure levels at a resolution of 2.5\degree, see][]{Kalnay1996}, the NCEP/DOE reanalysis II \citep{Kanamitsu2002}, ERA-40 \citep{Uppala2005} are often used and can be inputed in AtmoSwing. Other observed predictors can also be used, such as Sea Surface Temperature \citep[SST, ][]{Reynolds2007}. The predictors describing the target situation in operational forecasting can for example result from GFS \citep[Global Forecast System,][]{Kanamitsu1991,Kanamitsu1989}, which is operated by NCEP and NOAA.

\citet{Bontron2004} determined that the length of the archive should be 30 years to forecast usual situations, and 40 years for intense events. He also highlighted the dependence between the optimal number of analogues and the archive size: the shorter the archive, the lower the number of analogues. By varying the resolution of the predictors grids, he concluded that the performance increase is not significant for the analogy of atmospheric circulation below a resolution of 5\textdegree. However, it seems more important for the second level of analogy on moisture, which are more local variables. \citet{BenDaoud2010} assessed the European reanalysis ERA-40 \citep{Uppala2005} with a resolution of 1.125\textdegree. He identified a slight performance improvement, but not substantial enough to really justify an archive change \citep{BenDaoud2008}.

The predictand is the local meteorological variable of interest. It can be 6-houry data, or more often daily time series, but has to be consistent with the time step of the predictors (at least, not finer). The most used predictor is the daily precipitation, usually averaged over subregions in order to smooth local effects \citep{Obled2002, Marty2012}. These time series are frequently normalized by the precipitation value for a return period of 10~years \citep{Djerboua2001}. This normalization allows for an easier comparison between subregions subject to different precipitation regimes, and thus to better identify the most important contributions. Most users also apply a root square to this last ratio in order to reduce the weight of the extreme predictand values, but we generally don't.

\subsection{Present structure}

Although multiple variations of the AM exist, they generally share the same basics and the same structure. We want here to forecast the daily precipitation (the predictand) for a target day (at a defined lead time) and have long archives of both the predictand and predictor variables. The generic structure is the following:

\begin{enumerate}
	\item Preselection: to cope with seasonal effects, candidate dates are extracted from the archive within a period of four months centered around the target date, for every year of the archive. Alternatively, the candidate dates can be selected based on similar air temperature \citep{BenDaoud2010}.

	\item First level of analogy: we subsample $N_{1}$ dates within all the candidates provided by the preselection, by means of an analogy ranking. The first level of analogy is always the atmospheric circulation when it comes to precipitation forecasting. We thus assess the similarity of the atmospheric circulation of our target date with all the situations provided by the preselection by means of the S1 criteria \citep[Eq.\ref{eq:S1}, ][see also section \ref{sec:method:evolution}]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a certain spatial window.

	\begin{equation}
	\label{eq:S1}
	S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
	{\displaystyle \sum_{i} max\left\lbrace \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert \right\rbrace }
	\end{equation}
	where $\Delta \hat{z}_{i}$ is the forecast geopotential height difference between the \textit{i}th pair of adjacent points in the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both directions. The smaller the S1 values are, the more similar the pressure fields.

	The $N_{1}$ dates with the lowest values of S1 are considered as analogues to the target day. The number of analogues, $N_{1}$, is a parameter to calibrate.

	\item Subsequent level(s) of analogy: we can then subsample once again the $N_{1}$ analogues on the basis of another variable in order to obtain a lower number of analogue dates, $N_{2}$. The S1 criterion is usually not relevant for other predictors that the atmospheric circulation. Other classic criteria represent absolute distances: Mean Absolute Error (MAE) and Root Mean Squared Error (RMSE), the latest being most often used. These are more relevant to compare moisture variables for example.

	This process can be repeated, by subsampling a decreasing number of analogues, $N_{i}$, according to a ranking processed on various meteorological variables.

	\item Probabilistic forecast: then, the daily observed precipitation amount (or another predictand of interest) of the $N_{i}$ resulting dates provide the empirical conditional distribution considered as the probabilistic forecast for the target day. The empirical frequencies are processed for every value of the predictand, after classification, based on the Gringorten parameters (for a Gumbel or exponential law).

\end{enumerate}

\subsection{Proposed nomenclature}

Variants of the AM parameterization are numerous and it is not always easy to reference them in a short and descriptive way. We thus propose a basic nomenclature (Figure \ref{figure:nomenclature}) in order to give an idea of the structure in a simple identifier. This one cannot describe all the parameters of the AM, but quickly illustrate the structure of the implementation. This is particularly useful when working with a global optimization method, where nothing is fixed but the very structure of the AM.

\begin{figure}[htbp]
	\includegraphics[width=8.3cm]{figures/nomenclature.pdf}
	\caption{Proposed structure for naming the parameterizations of the AM.}
	\label{figure:nomenclature}
\end{figure}

The naming contains different blocs (separated by an hyphen) for the various levels of analogy. It starts with the specification of the preselection (P; can be omitted when comparing AMs with the same preselection approaches), which can be nowadays of 2 types:
\begin{itemize}
	\setlength\itemsep{-2px}
	\item C: calendar period ($\pm 60$ days around the target date)
	\item T: based on air temperature \citep{BenDaoud2010}
\end{itemize}

Then, the following levels of analogy are listed. They may start with an optional A (for analogy) when other methods are mixed into the parameterization \citep[see eg. ][]{Chardon2014}. For every level of analogy, we first provide the number of variables used (combination of atmospheric levels and time of observation) and then give the short name of the variable (according eg to ECMWF conventions; in upper case), for example:
\begin{itemize}
	\setlength\itemsep{-2px}
	\item Z : geopotential (circulation)
	\item MI : moisture index (TPW * RH)
	\item MF : moisture flux (V * TPW * RH)
	\item TPW : total precipitable water
	\item V : wind velocity
	\item W : vertical velocity
\end{itemize}

In order to keep the identifier simple, no value of atmospheric level neither time of observation is specified. Moreover, the analogy criterion is not specified and is supposed to be S1 for Z and RMSE for the other variables. If anything changes from these conventions, it can be noted as a flag. The flag (lower case) can also give other information, such as the optimization method:
\begin{itemize}
	\setlength\itemsep{-2px}
	\item cc : classic calibration (can be omitted as considered as default, see section \ref{sec:classic})
	\item go : global optimization (by means of genetic algorithms for example, see section \ref{sec:gas})
\end{itemize}

This nomenclature can be adapted to specific needs, or simplified for a better readability (eg. by removing the specification of the preselection). Examples are provided in the following section.

\subsection{Evolution of the method and examples}
\label{sec:method:evolution}

The evolution of the AM for precipitation forecasting led to several improvements. However, some of these are specific for a certain region and are not relevant for others, and some perform better depending on the lead time. Thereby, we cannot provide a single best parameterization for the AM, neither rank them from worst to best. We will thus provide some examples of parameterizations of the most recent AM variant, in the chronological order they appeared. The most common usage of the AM is for precipitation forecasting. Thus, the evolution presented hereafter focuses on this predictand. However, the AM, or an equivalent, was also used for short to medium term forecasting of daily temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983}, wind \citep{Gordon1987}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectory of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}. \citet{Guilbaud1997} performed a literature review about the use of the AM in long-term forecasting and identified operational applications for monthly forecasts in many countries, including Canada \citep{Shabbar1986},  Hungary \citep{Toth1989}, the Netherlands \citep{Nap1981}, and England \citep{Murray1974}, as well as seasonal forecasts: \citet{Barnett1978}, \citet{Bergen1982} and \citet{Livezey1988}.

\citet{Lorenz1969} introduced the concept and established the basis of the AM. He considered looking for analogues on the geopotential field (at 200, 500 and 850~hPa) with a corrected Euclidean distance averaged on the 3 levels. In addition, the analogue situations were to belong to the same period of the year to be comparable in terms of distribution of solar energy.

The use of the AM for operational forecasting of daily precipitation originates in the work of \citet{Duband1970, Duband1974, Duband1981}. It was then designed for operational forecasting at EDF (\'{E}lectricit\'{e} de France) in order to better manage water resources and flood risks. The geopotential height at 700~hPa and 00~h was used for its higher stability than the sea level pressure (SLP) field, and its sensitivity to the atmospheric disturbances \citep{Guilbaud1997}. However, the SLP at 06~h was also used for determining the intensity of precipitation, as well as the temperature at 700~hPa as indicator of the thermal state of the air mass \citep{Duband1974}. The data were then based on 37 radiosounding over Europe and condensed by a principal component analysis. The first 25 analogues on the geopotential at 700~hPa were considered, and multiple local regressions between precipitation, SLP, and the temperature at 700~hPa were established on these analogues before being applied to the target situation \citep{Guilbaud1997}. The archives were split by season to consider comparable situations in terms of distribution of solar energy \citep{Lorenz1969}. This initial rigid division into seasons was then transformed into a moving selection of more or less two months around the target date. The forecast was then performed only on the basis of observations and was temporally extrapolated to the following two days. After a few years, this statistical temporal extrapolation was abandoned in favor of a statistical adaptation using predictors resulting from an global numerical model, allowing a forecast over the following 4~days.

Thereafter, the geopotential was considered at 1000~hPa and 700~hPa and the analogues selection was done in two stages: first according to an Euclidean distance in the space of the principal components of the geopotential field at 700~hPa, and then according to a correlation criterion (on the geopotential field at 700 and 1000~hPa, as well as on the thickness between these two levels) in order to remove days too dissimilar \citep{Guilbaud1997}. The regression approach to determining precipitation was eventually abandoned in favor of the probabilistic forecasting (synthesized by the percentiles 20, 60 and 90~\%), as it is performed today by interpolating linearly over the cumulative empirical distribution of daily precipitation measured at the analogue days \citep{Guilbaud1997, Guilbaud1998}.

\citet{Wilson1980} sought the 20 best analogues on the geopotential field at 500 and 1000~hPa using the Teweles-Wobus (S1) criterion (Eq. \ref{eq:S1}), which allows for a comparison of the gradients and thus an analogy of the atmospheric circulation. \citet{Woodcock1980} used the same criterion on the SLP fields to select 50~analogues in order to forecast maximum daily temperatures. \citet{Yacowar1975} considered the 20 best analogues on the maximum of the sum of correlations on the two fields. This work has demonstrated the best performance of the AM compared to regression methods, as well as the interest to search for analogues on two (500 and 1000~hPa) geopotential fields combined \citep{Guilbaud1997}.

\citet{Mandon1985} introduced a second level of analogy and assessed different variables, such as wind, moisture, air temperature, surface temperature of the Mediterranean, the product of wind and the 700~hPa moisture, as well as a second barometric criteria differentiated according to the season. \citet{Vallee1986} continued to work on the second level of analogy and considered the wind at 700~hPa \& 12~h in order to improve heavy precipitation events for a region in France susceptible to southerly flows.

Taking advantage of higher-performance IT tools \citet{Guilbaud1997} stopped using PCA in order to work directly with the raw data interpolated on grids, which was an improvement, especially when considering the S1 criterion (already used for analogues selection by\citealp{Wilson1980} and \citealp{Woodcock1980}) instead of the Euclidean distance. This score gives importance to the similarity of atmospheric circulation rather than absolute values of the pressure field. After comparing this criterion to other criteria from the literature, or combinations of these, \citet{Guilbaud1998} found that S1 is the most efficient. They also pointed out the benefit of using 2 atmospheric levels, 1000~hPa (Z1000) and 700~hPa (Z700) or 500~hPa (Z500), instead of one (also instead of the thickness), and to also consider 2 temporal windows, 00~h and 24~h \citep{Obled2002}. The number of analogues considered is 50 (Table \ref{table:R0}). When using multiple predictors, they combined the criteria values calculated on each atmospheric level and each temporal window using an arithmetic mean. Finally, they also suggested using a second level of analogy on lower layers moisture and introduced the Ranked Probability Score \citep{Epstein1969} to assess the forecast quality, which was found to be more relevant that the Brier score \citep{Brier1950}.

\begin{table}[htbp]
	\caption{Method of type PC-4Z proposed by \citet{Guilbaud1997}.}
	\begin{tabular}{ccccc}
		\tophline
		\textbf{Level} & \textbf{Variable} & \textbf{Hour} & \textbf{Criteria} & \textbf{Nb} \\ 
		\middlehline
		0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
		\middlehline 
		\multirow{2}{*}{1} & Z1000 & 00, 24~h & \multirow{2}{*}{S1} & \multirow{2}{*}{50} \\
		& Z700 $^{\dagger}$ & 00, 24~h & & \\ 
		\bottomhline 
	\end{tabular} 
	\belowtable{$\dagger$ or Z500 as an alternative}
	\label{table:R0}
\end{table}

\citet{Bontron2004} could benefit from the newly available NCEP/NCAR reanalysis data. This dataset being more homogeneous, much longer, and contained more variables than what was previously used, he could systematically assess a large selection of variables in the AM combined with different criteria. He thus confirmed the interest of the geopotential height along with the S1 criterion for the first level of analogy. He has also shown that the choice of the temporal window (time of observation) has a higher importance than the atmospheric level for the performance of the AM. By assessing multiple combinations of atmospheric levels and temporal windows, he concluded that the coupled geopotential heights at 1000~hPa (Z1000) \& 12~h and 500~hPa (Z500) \& 24~h provided the best performance for the studied region (Table \ref{table:R1}). The analogy on the atmospheric circulation proposed by \citet{Bontron2004} is still, at the time of writing, often used operationally, and considered as a reference for benchmarking new parameterizations \citep{Horton2012, Daoud2015}.

\begin{table}[htbp]
	\caption{Method of type PC-2Z proposed by \citet{Bontron2004}.}
	\begin{tabular}{ccccc}
		\tophline
		\textbf{Level} & \textbf{Variable} & \textbf{Hour} & \textbf{Criteria} & \textbf{Nb} \\ 
		\middlehline 
		0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
		\middlehline 
		\multirow{2}{*}{1} & Z1000 & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{50} \\
		& Z500 & 24~h & & \\ 
		\bottomhline 
	\end{tabular} 
	\label{table:R1}
\end{table}

\citep{Bontron2004} also assessed several variables for the second level of analogy and identified the moisture variables (precipitable water and relative humidity) and vertical velocity to be the most relevant. Finally, he found that a moisture index made of the product of relative humidity at 850~hPa (RH850) and total precipitable water (TPW) gave the best performance (Table \ref{table:R2}). This index does not represent an actual physical quantity, but expresses the water content of the column and its proximity to saturation.

\begin{table}[htbp]
	\caption{Method of type PC-2Z-2MI proposed by \citet{Bontron2004}.}
	\begin{tabular}{ccccc}
		\tophline
		\textbf{Level} & \textbf{Variable} & \textbf{Hour} & \textbf{Criteria} & \textbf{Nb} \\ 
		\middlehline 
		0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
		\middlehline 
		\multirow{2}{*}{1} & Z1000 & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{70} \\
		& Z500 & 24~h & & \\ 
		\middlehline
		2 & TPW * RH850 & 12, 24~h & RMSE & 30 \\
		\bottomhline 
	\end{tabular} 
	\label{table:R2}
\end{table}

He noted the relevance of defining different spatial windows for each region, rather than a single window to the entire territory. These windows are more able to represent the specific atmospheric influences for the region. In order to assess the forecast, he introduced the Continuous Ranked Probability Score \citep[CRPS, Eq. \ref{eq:CRPS}, ][]{Brown1974}, which discards the issue of determining classes as with the RPS previously in use.

\citet{Gibergans-Baguena2007} implemented the AM for daily precipitation forecasts in Catalonia. They considered the geopotential at 700 and 1000~hPa, and the thickness between these two levels, to search for analogy of circulation, and then applied a second selection on local data from radiosonde: stability indexes, precipitable water, temperature, etc.

\citet{Marty2010} tested other temporal windows for intraday application on the basis of a more comprehensive reanalysis dataset. He first proposed to change the hours of observation for both levels of analogy (both for 06~h and 18~h), and selected the 925~hPa level for the moisture analogy (Table \ref{table:R3a}). The selected hours and atmospheric levels were not available before.

\begin{table}[htbp]
	\caption{Method of type PC-2Z-2MI proposed by \citet{Marty2010}.}
	\begin{tabular}{ccccc}
		\tophline
		\textbf{Level} & \textbf{Variable} & \textbf{Hour} & \textbf{Criteria} & \textbf{Nb} \\ 
		\middlehline 
		0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
		\middlehline 
		\multirow{2}{*}{1} & Z1000 & 06~h & \multirow{2}{*}{S1} & \multirow{2}{*}{75} \\
		& Z500 & 18~h & & \\ 
		\middlehline
		2 & TPW * RH925 & 06, 18~h & RMSE & 30 \\
		\bottomhline 
	\end{tabular} 
	\label{table:R3a}
\end{table}

Then, \citet{Marty2010} changed the moisture index into the moisture flux by adding the wind velocity (V700 or V925) component in the multiplication. He considered this flux at 700~hPa or 925~hPa (Table \ref{table:R3b}).

\begin{table}[htbp]
	\caption{Method of type PC-2Z-2MF proposed by \citet{Marty2010}.}
	\begin{tabular}{ccccc}
		\tophline
		\textbf{Level} & \textbf{Variable} & \textbf{Hour} & \textbf{Criteria} & \textbf{Nb} \\ 
		\middlehline 
		0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
		\middlehline 
		\multirow{2}{*}{1} & Z1000 & 06~h & \multirow{2}{*}{S1} & \multirow{2}{*}{60} \\
		& Z500 & 18~h & & \\ 
		\middlehline
		2 & V700 * TPW * RH700 $^{\dagger}$ & 06, 18~h & RMSE & 25 \\
		\bottomhline 
	\end{tabular} 
	\belowtable{$\dagger$ or V925 * TPW * RH925 as an alternative}
	\label{table:R3b}
\end{table}

\citet{BenDaoud2010} applied the AM in the context of large floodplains, out of an Alpine environment (Sa\^{o}ne, Seine). He evaluated several variables usually used in weather forecasting, amongst them the air temperature and the vertical velocity that presented a potential interest. The temperature (at the nearest grid point, on the 925~hPa (36~h) and 600~hPa (12~h) levels) could replace the rigid calendar preselection of $\pm$ 60 days around the target date by a more dynamic screening of similar situations in terms of air masses. The seasonal effect is indeed also present in the temperature data. The number of preselected dates is equivalent to the number of days we would have chosen with the calendar approach, and thus depends on the archive size: $N_{0} = 60 \cdot 2 \cdot n_{a}$ where $n_{a}$ is the number of years in the calibration period. \citet{BenDaoud2010} also reconsidered the parameters of the moisture index and ended up with both 925~hPa and 700~hPa levels, at 12~h and 24~h, or at every time step (6~h) between 06~h and 30~h (Table \ref{table:R4b}).

\begin{table}[htbp]
	\caption{Method of type P2T-2Z-4MI (or P2T-2Z-10MI for the alternative version), by \citet{BenDaoud2010}.}
	\begin{tabular}{ccccc}
		\tophline
		\textbf{Level} & \textbf{Variable} & \textbf{Hour} & \textbf{Criteria} & \textbf{Nb} \\ 
		\middlehline
		\multirow{2}{*}{0} & T925 & 36~h & \multirow{2}{*}{RMSE} & \multirow{2}{*}{$N_{0}$} \\
		& T600 & 12~h & & \\ 
		\middlehline 
		\multirow{2}{*}{1} & Z1000 & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{70} \\
		& Z500 & 24~h & & \\ 
		\middlehline
		\multirow{2}{*}{2} & TPW * RH925 & 12, 24~h $^{\dagger}$ & \multirow{2}{*}{RMSE} & \multirow{2}{*}{25} \\
		& TPW * RH700 & 12, 24~h $^{\dagger}$ & & \\ 
		\bottomhline 
	\end{tabular} 
	\belowtable{$\dagger$ or 06-30~h as an alternative}
	\label{table:R4b}
\end{table}

Subsequently, \citet{BenDaoud2010} added an additionally level of analogy, between the circulation and the moisture analogy (Table \ref{table:R5}), based on vertical velocity at 850~hPa (W850). This AM was developed for large plains in France, but was found to be not relevant for an Alpine environment where topography is mostly responsible for uplift of air masses \citep{Horton2012}.

\begin{table}[htbp]
	\caption{Method of type P2T-2Z-4W-4MI, by \citet{BenDaoud2010}.}
	\begin{tabular}{ccccc}
		\tophline
		\textbf{Level} & \textbf{Variable} & \textbf{Hour} & \textbf{Criteria} & \textbf{Nb} \\ 
		\middlehline
		\multirow{2}{*}{0} & T925 (1 point) & 36~h & \multirow{2}{*}{RMSE} & \multirow{2}{*}{$N_{0}$} \\
		& T600 (1 point) & 12~h & & \\ 
		\middlehline 
		\multirow{2}{*}{1} & Z1000 & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{170} \\
		& Z500 & 24~h & & \\ 
		\middlehline
		2 & W850 & 06-24~h & RMSE & 70 \\
		\middlehline
		\multirow{2}{*}{3} & TPW * RH925 & 12, 24~h & \multirow{2}{*}{RMSE} & \multirow{2}{*}{25} \\
		& TPW * RH700 & 12, 24~h & & \\
		\bottomhline 
	\end{tabular}
	\label{table:R5}
\end{table}

\citet{Bliefernicht2010} proposed to add a weighting to the grid points of the predictor to overcome a comparison on a fixed-size window, and to give more importance to regions with relevant atmospheric circulation information. However, he noticed that there is a risk of overparametrization and a simplification of the weighting fields may be necessary, and that the performance was not stable between the calibration and the validation periods. He also compare the performance of the AM with a forecasting method based on a weather types approach, and concluded that the AM was more efficient.

\citet{Horton2012} identified that the analogy on the atmospheric circulation may not be comprehensive, and that a significant gain was obtained by considering 4 atmospheric levels (1000, 850, 700, and 500~hPa at 12 or 24~h) instead of 2 (Table \ref{table:R6}). No number of analogues ($N_{1}$) was provided globally, as it was calibrated for every subregion. Considering a AM only based on the atmospheric circulation has an interest for operational forecasting, where moisture variables are less well forecasted over a lead time of a few days. An equivalent structure was also optimized by means of generic algorithms (PC-4Zgo), which is the topic of another paper.

\begin{table}[htbp]
	\caption{Method of type PC-4Z proposed by \citet{Horton2012}.}
	\begin{tabular}{ccccc}
		\tophline
		\textbf{Level} & \textbf{Variable} & \textbf{Hour} & \textbf{Criteria} & \textbf{Nb} \\ 
		\middlehline 
		0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
		\middlehline 
		\multirow{4}{*}{1} & Z1000 & 12~h & \multirow{4}{*}{S1} & \multirow{4}{*}{$N_{1}$} \\
		& Z850 & 24~h & & \\ 
		& Z700 & 12~h & & \\ 
		& Z500 & 24~h & & \\ 
		\bottomhline 
	\end{tabular}
	\label{table:R6}
\end{table}

\citet{Horton2012} also integrated an analogy on the moisture flux, after the circulation analogy on the 4 atmospheric levels (Table \ref{table:R7}). The number of analogues ($N_{1}$, $N_{2}$) were also calibrated for every subregion. Similarly, an equivalent structure was optimized (PC-4Zgo-2MIgo).

\begin{table}[htbp]
	\caption{Method of type PC-4Z-2MI, by \citet{Horton2012}.}
	\begin{tabular}{ccccc}
		\tophline
		\textbf{Level} & \textbf{Variable} & \textbf{Hour} & \textbf{Criteria} & \textbf{Nb} \\ 
		\middlehline 
		0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
		\middlehline 
		\multirow{4}{*}{1} & Z1000 & 12~h & \multirow{4}{*}{S1} & \multirow{4}{*}{$N_{1}$} \\
		& Z850 & 24~h & & \\ 
		& Z700 & 12~h & & \\ 
		& Z500 & 24~h & & \\ 
		\middlehline
		2 & V700 * TPW * RH700 & 12, 24~h & RMSE & $N_{2}$ \\
		\bottomhline 
	\end{tabular}
	\label{table:R7}
\end{table}

% TODO : Sabine
% TODO : Jérémie Chardon

\subsection{Discussion on the method}

A version of the AM was evaluated during the project STARDEX \citep[\textit{STAtistical and Regional dynamical Downscaling of EXtremes for European regions}, see][]{Goodess2003, Stardex2005}. One of the project goals was to compare various downscaling methods for the determination of weather extremes, and the AM was selected among the most interesting from various techniques \citep{Maheras2005, Schmidli2007}. Its use as adaptation method of climate models is also the subject of a growing number of studies \citep{Zorita1999, Wetterhall2005, Wetterhall2007, Matulla2007, Chardon2014, Dayon2015}.

\citet{Hamill2006} used an analogy based approach on the GFS reforecasts in order to correct systematic errors in the ensemble forecasts of temperature and precipitation. Indeed, the statistical approach showed some bias in estimates of the numerical model, which may have been adjusted by taking into account the intrinsic local climatology from the AM. Moreover, the under-dispersion of the ensemble forecast from the numerical model has been corrected using analogues \citep{Hamill2006}. Correction of ensemble forecast under-dispersion by means of the AM is also used operationally at EDF.

\citet{Bliefernicht2010} observed that the performance of the AM is higher for winter than for summer. The relationship between synoptic predictors and local rainfall is lower in summer, due to convective precipitation that present a higher spatial variability and that depend on other parameters. Variables describing the synoptic circulation are indeed not able to predict the location of thunderstorm cells. This was also observed by \citet{BenDaoud2010}, who set up a specific model for the summer months (June 15 to September 15).

One of the AM limitations is the need for a substantial archive of the predictand variable, for example measured precipitations. Without data on several decades, the AM is not applicable. Long predictor archives are also required, but this problem is usually solved with reanalysis data, which are not perfect in terms of homogeneity, but can be considered of sufficient quality. Moreover,  reanalysis data are available all around the world, which represent a great potential for the AM. Another issue, still about data, concerns the operational forecasting: predictors describing forecasted target situations and the ones from the archive are not fully homogeneous, as they usually don't result from the same model. It is therefore necessary to use robust predictors that depend from the numerical model characteristics at a minimum extent.

Another limitation is the fact that extreme events may be under-represented in the considered sample of analogue dates. Indeed, in a limited weather archive, events with high return periods are not so numerous. Their number is certainly lower than the number of analogues we consider, which can introduce a bias in the forecast. There are however techniques to correct this bias \citep[see][]{Marty2010}.

It is legitimate to raise the question of the relevance of an approach based on archives of past situations in a context of climate change. The first potential issue is a possible change in atmospheric circulation. \citep{Philipp2007} observed certain trends in the synoptic atmospheric circulation over Europe, but with moderate frequencies. Moreover, the basic laws governing the atmosphere behavior will not be transformed \citep{Hewitson1996}. The assumption is that a large part of local climate change will result from changes in intensity, frequency and persistence of synoptic variables, but with other characteristics substantially similar to the present situation \citep{Hewitson1996}. Thus, if the archive of weather situations is long enough, it is reasonable to assume that a large part of future situations is already represented, even those whose frequency will change under different climatic conditions \citep{Wetterhall2005}. In addition, climate change is relatively slow, and through continuous updating of the archive, the AM will incorporate new information progressively with no big discontinuity.

\section{Software basics}

AtmoSwing is made of 3 main modules (Figure \ref{figure:flowchart_modules_atmoswing}) that are standalone, but do share a common code basis: the Forecaster, to process the operational forecasting, the Viewer, to display the forecast in a GIS environment, and the Optimizer, to calibrate/optimize the statistical relationship defining the analogy for a given predictand time series. Separating the Forecaster and the Viewer allows to automate the forecast on a server and to quickly display the results locally. The code is written in object oriented C++ and relies on the wxWidgets \citep{Smart2006} library to provide a cross-platform native experience to users. The software can thus be compiled on MS Windows, Linux / Unix, Mac (OS X, iOS), both in 32-bit and 64-bit.

\begin{figure}[htbp]
	\includegraphics[width=8.3cm]{figures/logo.pdf}
	\caption{The AtmoSwing logo and its 3 modules: the Forecaster, the Viewer and the Optimizer.}
	\label{figure:logo}
\end{figure}

The source code in under version control (with Mercurial), and is open source (on Bitbucket, www.atmoswing.org). Developments have been made with an approach of unit tests. A collection of 500+ tests are frequently evaluated and completed during development, in order to verify that code changes do not introduce regression. Every analogy criterion, prediction score, searching and sorting functions, data manipulation, and so on, is tested. The integration tests specific to the  AM rely on the results of another analogue sorting software developed in Grenoble. They result from a comparison work done with the help of R. Marty in order to ensure that the results of AtmoSwing are exactly equivalent to their model, given the same parameters and data. These tests are part of the checks carried out regularly, so that these results are always reproducible.

\subsection{Modular approach}

An AtmoSwing great strength is that it is designed to process analogy forecasting in a modular fashion. The structure of the AM (number of analogy levels, number of predictors) is built dynamically (Figure \ref{figure:flowchart_modules_atmoswing}), and nothing is fixed a priori. The software then performs successively as many analogy levels as the user specified, with all the predictors he wants. Each level of analogy processed results in an object containing target dates, analogue dates, values of the analogy criteria, and other data. This object can be saved in a NetCDF file and/or can be injected into a new analogy level. The whole structure of the AM is defined in an XML file generated by the user.

\begin{figure}[htbp]
	\includegraphics[width=8.3cm]{figures/flowchart_modules_atmoswing.pdf}
	\caption{Simplified organizational chart of the AM implementation in AtmoSwing.}
	\label{figure:flowchart_modules_atmoswing}
\end{figure}

Each implementation of the AM (see section \ref{sec:method:evolution}) may enter this scheme, even if it consists of manipulated variables (e.g. moisture index). Various preprocessing functions are implemented, as the calculation of the moisture index or flux, multiplication operations or calculation of the gradients. The user can specify the preprocessing method and the predictors to use, dynamically, in the XML file.

\subsection{Performance}

Although processing an analogue forecast for a given target date is fast, when optimizing the AM, forecasts are processed over a long periods of several decades, which is time consuming. Thus, great effort has been made in order to reduce the processing time to a minimum. Waste of time were identified with profiling tools and cut down at different levels. The software also use the linear algebra library Eigen 3 \citep{Guennebaud2010} for calculations on vectors and matrices, which resulted in time saving. Obviously, multi-threading is also implemented.

First, every identified redundancy has been removed. Then, when looking for a certain date or data, the search looks first in the proximity where it is likely to be found instead of exploring an entire array. Similar data are not loaded twice, but shared pointers are used. Many other improvements allowed saving time, for example the use of the quicksort method \citep{Hoare1962a} to sort the date vectors according to the values of analogy criteria. Different implementations were tested (many remain in the code as alternatives) in order to select the most efficient: for example, when storing analogue dates according to their criteria value, it is faster to insert them in a fixed-size array rather than storing them all and sorting the array subsequently.

The major performance gains were obtained by:
\vspace*{-2mm}
\begin{itemize}
	\setlength\itemsep{-2px}
	\item reduction of the array sizes used in the calculations,
	\item use of pointers (and therefore the decrease of data copies),
	\item better search for dates in temporal vectors starting from the previous index,
	\item better management of the analogue vectors,
	\item preprocessed gradients on geopotential fields when using the S1 criteria,
	\item and finally the implementation of parallel calculations (multi-threading).
\end{itemize}

\section{AtmoSwing Forecaster}

The Forecaster module allows to process operational forecasts. This software processes a current or past forecasting, defined by an XML file. A list of parameterizations (defined by external XML files) is processed successively. The software can compiled with a GUI (), or without in order to be used on a headless server through a command line interface. Processing a forecast requires very low computing capabilities and can be done on a low-end computer.

\begin{figure}[htbp]
	\includegraphics[width=8.3cm]{figures/atmoswing-forecaster.png}
	\caption{Graphical user interface of the Forecaster module. A list of different parameterizations is being processed.}
	\label{figure:atmoswing-forecaster-gui}
\end{figure}

The software first downloads the predictor describing the target situation, such as GFS outputs \citep[Global Forecast System,][see section \ref{sec:data}]{Kanamitsu1991,Kanamitsu1989}. It then interpolates the gridded data to match the resolution of the archive (for example 2.5\textdegree\ in the case of the NCEP/NCAP reanalysis I). The analogues dates are next sought according to the provided parameterization, and the predictand data are associated with the corresponding dates. The results are finally saved in auto-describing NetCDF files. If required, a synthetic XML is generated for an easier integration on a web platform for example. Every step of the forecast, from the predictor downloading to the final results, is done in the software (and controlled through configuration), without use of external scripts (e.g. for data conversion).

The command-line interface allows to start the process for the most recent date, a selected past date, or the last $x$ days (as long as the predictors are available on the provider server). When there is nothing to process (no new predictor data available), the execution just stops, leading to no loss of computing resources. The recommended use is thus to set up a cron task on a Linux server in order to trigger the forecast every 30 minutes. When using GFS outputs, this would provide 4 forecasts a day with a reduced delay between GFS outputs availability and the analogue forecast.

A user interface allows the creation of the precipitation database. Its generation consists in extracting the time series from text files in order to process them and save a database in the NetCDF format. During the process, Gumbel adjustments are automatically calculated in order to determine the values of different return periods. The time series are normalized by a selected return period (default 10~years) and their square root is eventually processed. The final database file contains both the raw and the normalized series, as well as characteristics of the gauging stations and some metadata.

\section{AtmoSwing Viewer}

\subsection{The graphical user interface }

The Viewer module allows displaying the resulting forecast files in an interactive GIS environment (Figure \ref{figure:atmoswing-viewer-gui}). During the forecast, different parametrization variants of the same type of AM may be applied to a region; the reason being that some parameterizations may be specific for a subregion, when its meteorological influences differ, which is often the case even in relatively small catchments \citep{Horton2012}. The Viewer automatically gathers the similar AMs and provides a composite view of the optimal forecasts per subregion. The user can however choose to display a specific parameterization for the whole region, which is more consistent in terms of homogeneity of the analogue dates.

\begin{figure*}[t]
	\includegraphics[width=18cm]{figures/atmoswing-viewer.png}
	\caption{Graphical user interface of the Viewer module.}
	\label{figure:atmoswing-viewer-gui}
\end{figure*}

The software provides several levels of synthesis of the forecasts. When it is opened, it automatically loads the most recent forecasts. It first offers a quick overview of possible alerts by means of color codes on the lead time switcher (upper right in the GUI, see Figure \ref{figure:atmoswing-viewer-gui}), which synthesizes the worst case scenario, or in the alarms panel (on the left part of the GUI). The alarms panel offers a synthesis of the higher forecasted values over the region, for the different AMs and the different lead times. By default, the colors are expressed relatively to 10-year return period values, for the $90^{th}$ percentile (can be changed in the preferences). This higher level of synthesis allows to quickly identify potential critical situations in the days ahead.

Then, the user can explore the forecasts more in details, starting from the map provided by the GUI (Figure \ref{figure:atmoswing-viewer-gui}). The map display the forecast of the selected parameterization (upper left) and the selected lead time (upper right). The displayed AM is by default a composite of the parameterization variants for the region, but the user can choose a specific parameterization for the whole region by opening the tree view and selecting a child element. We also allowed a display of all lead times on a single map by means of a symbolic representation on a circular band with a box for every lead time (Figure \ref{figure:atmoswing-viewer-snail}). The number of boxes is dynamically adjusted to the number of lead times. This representation offers a global spatiotemporal display for a chosen parameterization.

Color scales in the map are easily controlled by choosing the predictand ratio (raw value or ratio to different return periods) and the quantile of the distribution (on the left part of the GUI). Representation relatively to a return time better expresses the meaning of a certain precipitation amount for a given region, as climatology can be drastically different between two locations in a mountainous area, even if they are close to each other. All information relative to a rain gauge station (or stations groups), such as its location, its name, or the values of different return periods, are stored in the forecast files, in order to be displayable for end users who do not have the predictand database.

\begin{figure}[htbp]
	\includegraphics[width=8.3cm]{figures/atmoswing-viewer-snail.png}
	\caption{Visualization of multiple lead times on the map.}
	\label{figure:atmoswing-viewer-snail}
\end{figure}

By clicking on a station (forecast points) on the map (or by selection in a dropdown list on the left), a new window appears with a plot of the forecasted time series (Figure \ref{figure:atmoswing-viewer-timeseries}). By default, the plot contains the usual 3 considered percentiles (90$^{th}$, 60$^{th}$, and 20$^{th}$), along with the 10 best analogues with a color code from yellow (10$^{th}$) to red (1$^{st}$). The 10 year return period is also displayed in order to put the forecast in perspective. The user can choose to hide any data or to display supplementary information (all analogues, all 10$^{th}$ percentiles, or all return periods) in the left panel. Traces of old forecasts are also automatically loaded and displayed in order to inform on the stability of the forecasts.

\begin{figure}[htbp]
	\includegraphics[width=8.3cm]{figures/atmoswing-viewer-timeseries.png}
	\caption{Visualization of the forecasted time series.}
	\label{figure:atmoswing-viewer-timeseries}
\end{figure}

\begin{figure}[htbp]
	\includegraphics[width=8.3cm]{figures/atmoswing-viewer-distribution.png}
	\caption{Visualization of the forecasted precipitation distribution for a given lead time.}
	\label{figure:atmoswing-viewer-distribution}
\end{figure}

The user can even go into further details and display the predictand cumulative distribution for a given lead time (Figure \ref{figure:atmoswing-viewer-distribution}). This information is useful when working on extreme precipitation, as comparing the distribution of all analogues versus the 10 best analogues provides information in order to interpret the forecast (trend to under/overestimation). Distribution of the analogy criteria can also be displayed in order to identify eventual discontinuities in the criteria values.

Finally, one can see the details of the analogue dates, along with the predictand and the criteria values in an interactive spreadsheet.

AtmoSwing Viewer relies on workspaces to specify the path to the forecast directories and the GIS layers. Many GIS formats are supported thanks to the GDAL and VroomGIS libraries. One can have as many layers as desired, and can control their display properties. It is thus easy to switch from a forecast for a region to another by opening a workspace file (XML format).

\subsection{Interpreting a forecast}

\citet{Djerboua2001} applied the AM developed by \citet{Guilbaud1997} in operational forecasting as part of the MAP project \citep[\textit{Mesoscale Alpine Programme}, see][]{Binder1996} during the special observation period \citep{Bougeault2001}. He noted that the 60$^{th}$ percentile was better to forecast precipitation amounts for common situations, but for strong to extreme events, the 90$^{th}$ percentile turned out to be a better indicator. It is therefore necessary to pay particular attention to the 90$^{th}$ percentile which, when reaching high values, may be indicative of a situation related to extreme precipitations, due the presence of several analogue dates with significant observed precipitations amounts in the distribution \citep{Djerboua2001}. \citet{Bontron2004} made the same observation: he found the 60$^{th}$ percentile to be best for occurrence forecasting, but the 90$^{th}$ percentile to be the most informative in order to forecast medium to high precipitation amounts ($P > P10/10$, one tenth of the 10-year return period value). In this regard, operators should pay particular attention to the distribution of the best analogues. If, for example, it is shifted to higher values, there is a risk of under-estimating the event when considering the middle of the full distribution.

When comparing the performance of Meteo France numerical models along with the AM, \citet{Djerboua2001} concluded that the numerical models are better at predicting the current day, but the AM becomes more competitive for the following lead times. It is indeed very interesting in early warning approaches, as it is able to identify a potentially critical situation one week ahead.

All AMs rely on atmospheric circulation information, while some add subsequent levels of analogy on moisture or other thermodynamic predictors (see section \ref{sec:method:evolution}). However, these subsequent predictors were found to be relevant only on the first 3~lead times, and not much further, due to higher uncertainties in their forecast. This was found to be the case of moisture indexes and vertical velocity. The user has to keep this aspect in mind when interpreting forecasts over a range of lead times.

\section{AtmoSwing Optimizer}

AtmoSwing Optimizer is a single tool that integrates different optimization methods, presented in section \ref{sec:atmoswing-calibration} to \ref{sec:atmoswing-optimization}.

\subsection{Calibration framework}

The calibration of the AM is usually carried out in a perfect prognosis \citep{Klein1959} framework \citep{BenDaoud2010, Bontron2004}. Perfect prognosis uses observed or reanalyzed data to calibrate the relationship between predictors and predictands. Then, when used in operational forecasting, this relationship is applied to global model forecasts, that contains larger uncertainties. This framework allow us to identify relationships that are as close as possible to the natural links between predictors and predictands, by reducing uncertainties related to numerical forecasting models. However, no model is perfect, and even reanalysis data contain a bias that cannot be ignored. For this reason, the statistical relationships identified in the perfect prognosis framework should be applied to model outputs that are as similar as possible to the model used to elaborate the reanalysis.

Another reason for working in a perfect prognosis framework is that numerical models evolve continuously, and so does the forecast they provide. Re-forecasts would allow us to work on a homogeneous dataset, as they are regularly reprocessed. However, we would need to redo the calibration procedure every time a new version is available, in order to reduce the bias \citep{Wilson2002}. Moreover, the reforecasts are not re-processed for every new version of the model, meaning we still end up with a bias between the forecast and the archive. Finally, the length of reanalyses datasets are usually much longer than reforecast datasets, which allows us to identify more robust relationships. The size of the archive is indeed an important criteria for the AM.

The statistical relationship is established on a calibration period that is as long as possible. For every day of this period, a search for analogues is processed, the precipitation data are associated with the corresponding dates and a forecast score is calculated. During the search for analogues situations, 120 days around the target date are excluded (thus excluding data in the same year) in order to consider only truly independent candidates days.

A validation period is always considered. It consists of an independent period that is never used as target neither candidate date. Validating the parameters of the AM is very important in order to avoid over-parametrization and thus to ensure that the statistical relationship is valid on another period. In order to account for climate change and the evolution of the measuring techniques, it is recommended to use a noncontinuous period for validating, distributed over the whole archive \citep{BenDaoud2010}. The user can thus specify a list of the years to keep for the validation and AtmoSwing Optimizer handles this aspect easily by both removing the validation years from possible candidate situations, and by automatically processing the validation score at the end of the optimization.

\subsection{Implemented forecasts scores}

Most often, the accuracy of the parameters for precipitations forecasting is evaluated by means of the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It is the only score we detail here (in section \ref{sec:prob_forecasts}). In addition, other scores are implemented in AtmoSwing Optimizer. The following sections list the scores presently available, but will not detail them.

\subsubsection{Discrete deterministic forecasts}

Deterministic forecasts consider no uncertainty. A single value is provided for a variable. The first group of this type of forecast is based on discrete values, also called categories. These scores are used, for example, to evaluate a deterministic prediction of threshold exceedance. In our case, the continuous probabilistic nature of an analogy forecast can be transformed into a discrete forecast by considering a threshold exceedance and a fixed percentile from the distribution. These scores rely on a contingency table, which is here a table containing the sums of observed / unobserved and predicted / unpredicted events \citep{Wilks2006}. On the basis of this table, multiple scores can be processed within AtmoSwing:

\begin{itemize}
	\setlength\itemsep{-1mm}
	\item Proportion correct \citep{Finley1884}
	\item Threat Score \citep{Gilbert1884}
	\item Bias
	\item False Alarm Ratio
	\item Hit Rate or Probabiliy of Detection
	\item False Alarm Rate
	\item Heidke Skill Score \citep{Heidke1926}
	\item Peirce Skill Score \citep{Peirce1884}
	\item Gilbert Skill Score or Equitable Threat Score \citep{Gilbert1884}
\end{itemize}

\subsubsection{Continuous deterministic forecasts}

The difference of the continuous deterministic forecasts compared to discrete ones is that they provide non-discrete values that must be evaluated with distance measures. In our application, the distribution of the analogues is summarized by a chosen percentile. Available scores are the following:

\begin{itemize}
	\setlength\itemsep{-1mm}
	\item Mean Absolute Error
	\item Root Mean Squared Error
\end{itemize}

\subsubsection{Discrete probabilistic forecasts}

In the case of discrete probabilistic forecasts, we work with probabilities of occurrence or probability of belonging to a certain category. The implemented options are:

\begin{itemize}
	\setlength\itemsep{-1mm}
	\item Brier Score \citep{Brier1950, Toth2003}
	\item ROC diagramm \citep[Relative Operating Characteristic or Receiver Operating Characteristic,][]{Mason1982}
	\item RPS \citep[Ranked Probability Score,][]{Epstein1969}
	\item SEEPS \citep[Stable Equitable Error in Probability Space,][]{Rodwell2010,Rodwell2011}
\end{itemize}

\subsubsection{Continuous probabilistic forecasts}
\label{sec:prob_forecasts}

Continuous probabilistic forecasts of a variable are issued in the form of the expected statistical distribution for this variable. They are expressed either in terms of distribution or by probability density, which need to compared an observed value. This is the situation we meet when forecasting precipitations with the AM.

We mainly rely on the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows assessing the predicted cumulative distribution functions $F(y)$ compared to the observed value $y^{0}$. The better the forecast, the smaller the score. The mean CRPS of a forecast series of length $n$ can be written:

\begin{equation}
\label{eq:CRPS}
CRPS = \frac{1}{n} \sum_{i=1}^{n} \left(  \int_{-\infty}^{+\infty} \left[ F_{i}(y)-H_{i}(y-y_{i}^{0})\right]^{2} dy \right) 
\end{equation}
where $H(y-y_{i}^{0})$ is the Heaviside function that is null when $y-y_{i}^{0}<0$, and has the value 1 otherwise. The mean CRPS is averaged on the calibration, respectively the validation periods, on all days, may they be dry, slightly rainy or with heavy precipitation.

This score is now commonly used for the evaluation of continuous variables prediction systems \citep{Casati2008, Marty2010}. To compare the value of the score in regard to a reference, we often consider its skill score value, and use the climatological distribution as the reference. The CRPSS (\textit{Continuous Ranked Probability Skill Score}) is thus defined as following:

\begin{equation}
\label{eq:CRPSS}
CRPSS = \dfrac{CRPS-CRPS_{r}}{CRPS_{p}-CRPS_{r}} = 1-\dfrac{CRPS}{CRPS_{r}}
\end{equation}
where $CRPS_{r}$ is the CRPS value for the reference and $CRPS_{p}$ would be the one for a perfect forecast ($CRPS_{p} = 0$).

The CRPS can be decomposed into several indicators, also implemented into AtmoSwing Optimizer, such as:

\begin{itemize}
	\setlength\itemsep{-1mm}
	\item Reliability \citep{Hersbach2000}
	\item Resolution / uncertainty \citep{Hersbach2000}
	\item Sharpness \citep{Bontron2004}
	\item Accuracy \citep{Bontron2004}
\end{itemize}

Finally, the rank diagram \citep{Talagrand1997} and its accuracy as defined by \citet{Candille2005} are also available.

\subsection{The sequential calibration}
\label{sec:atmoswing-calibration}

The calibration procedure that we call ''sequential'' or ''classic'', but which is also named ''optimization'' \cite[eg. by ][]{BenDaoud2015}, was developed by \citet{Bontron2004} at the LTHE laboratory (INPG, Grenoble). It is a semi-automatic procedure that determines the optimal parameters for the different variables of each level of analogy. The analogy levels (eg. the atmospheric circulation or moisture index) are calibrated sequentially. The procedure consists of the following steps \citep{Bontron2004}:

\begin{enumerate}
	\item Manual selection of the following parameters:
	\begin{enumerate}
		\item meteorological variable,
		\item pressure level,
		\item temporal window (hour of observation),
		\item initial analogue numbers.
	\end{enumerate}

	\item For every level of analogy:
	\begin{enumerate}
		\item Identification of the most skilled unitary cell (1~point for moisture variables and 4 for the geopotential height when using the S1 criteria) over a large domain. Every point (or cell) of the full domain is assessed jointly on all predictors of the level of analogy (consisting generally of the same variable, but on different pressure levels and at different hours).
		\item From this most skilled point, the spatial window is expanded by successive iterations in the direction of greater performance gain. The detailed stages are the followings: (i) The unitary spatial window is expanded in every 4 directions successively. The performance score is processed for these 4 windows. (ii) Only the direction providing the best improvement is applied to our spatial window. (iii) From this new spatial window, an increase in every 4 directions is once again assessed, and the best improvement is applied. (iv) The spatial window grows up by repeating the previous steps, until no improvement is reached.
		\item The number of analogues is optimized for the current level of analogy.
		\item A new level of analogy can be added, based on other variables (such as the moisture index) on predefined pressure levels and time frames. The number of analogues for the next level of analogy is initiated at a chosen value. Then, the procedure starts again from step (a) for the new level. The parameters calibrated on the previous analogue levels are fixed and do not change (except the number of analogues, at the final stage). 
	\end{enumerate}
	\item Finally, the numbers of analogues are re-assessed for the different levels of analogy. This is done iteratively by varying the number of analogues of each level in a systematic way.
\end{enumerate}

The calibration is thus done in successive steps, on a limited number of parameters. Previously calibrated parameters are generally not reassessed. The advantage of this method is that it is fast, it provides acceptable results, and it has low computing requirements.

We added small improvements to this method, and named it ''classic+'', by allowing the spatial windows to do other moves, such as: (1) increase in 2 simultaneous directions, (2) decrease in 1 or 2 simultaneous directions, (3) expansion or contraction (in every direction), (4) shift of the window (without resizing) in 8 directions (including diagonals), and finally (5) all the moves described above, but with a factor of 2, 3, or more. For example, we try to increase by 2 units in one (or more) direction. This allows to skip one size that may not be optimal. These supplementary steps often result in spatial windows that are a bit different, but the performance gain is rather marginal.

\subsection{Monte-Carlo analysis}

A Monte-Carlo analysis is also implemented in AtmoSwing. It performs thousands assessments of random parameters within given ranges. This method is not efficient to find the best parameters set, but helps to better understand the sensitivity of the parameters.

For example, a Monte-Carlo analysis was performed on the PC-2Z (Table \ref{table:R1}) AM for the southeastern crests on the upper Rh\^{o}ne catchment in Switzerland, with 10,000 random selections of its parameters according to a uniform law (the analysis was also performed in other subregions of the upper Rh\^{o}ne catchment in Switzerland, but the conclusions depicted here are similar). Unlike conventional calibration, the spatial windows on Z500 and Z1000 were not necessarily overlapping. Thus, 9 parameters were varying together (8 for the spatial windows and 1 for the number of analogues)

Figure \ref{figure:monte_carlo_r1} presents the resulting scatter of random exploration of the parameters space, along with the results of the sequential calibration (red cross). Results are illustrated here for the 4 parameters defining the spatial window on Z500. Despite the high number of simulations, the upper portion of the distributions seems incomplete, and the score of the sequential calibration is not matched. Although the method is poor to find efficiently the best parameters, it is informative as to the shape of the scatter plots.

The shape of the scatter plots from Figure \ref{figure:monte_carlo_r1} reveals a low sensitivity as to the exact location of the spatial windows. There is a maximum threshold for minimum longitudes and latitudes and, conversely, a minimum threshold for maximum latitudes and longitudes, but outside of these limits, the distribution slope is more gentle. A spatial window can thus be a bit larger than its optimal size without a significant drop in performance, as long as it includes at least one critical region. This was also observed by \citet{Bontron2004}, who noted that ''\textit{performance slowly decrease if we consider a slightly too large window, while the use of too small windows results in strong performance loss}''. The dilution of a part of the relevant synoptic information has therefore not necessarily a significant negative impact on performance, while ignoring some of this information leads to undesirable loss of performance. It is the same for the analogues number (not shown), for which, past an optimum, the upper slope of the distribution decreases slowly.

\begin{figure}[htbp]
	\includegraphics[width=8.3cm]{figures/monte_carlo_r1.png}
	\caption{Example of a Monte Carlo analysis on the PC-2Z (Table \ref{table:R1}) AM. Results are illustrated for the 4 parameters defining the spatial window on Z500, along with the results of the sequential calibration (red cross).}
	\label{figure:monte_carlo_r1}
\end{figure}

The same analysis has been conducted on the AM with 2 levels of analogy, PC-2Z-2MI (Table \ref{table:R2}), and the results on the number of analogues on the 1st and the 2nd level are illustrated in Figure \ref{figure:monte_carlo_r2}. One can see that the AM is not very sensitive to the number of analogues on the first level of analogy, when considering a second level afterwards, as long as this number is not too small. The trend on the second level is more apparent, but a large tolerance seems possible.

\begin{figure}[htbp]
	\includegraphics[width=8.3cm]{figures/monte_carlo_r2.png}
	\caption{Example of a Monte Carlo analysis on the PC-2Z-2MI (Table \ref{table:R2}) AM. Results are illustrated for the number of analogues on the 1st and 2nd level of analogy, along with the results of the sequential calibration (red cross).}
	\label{figure:monte_carlo_r2}
\end{figure}

The use of such a random method allows informing on the parameters sensitivity, and so on the importance of the expected precision of the calibration results. It helps understanding how different parameterizations can end up to relatively similar performances. Using Mote Carlo analyses allows to easily display the parameters sensitivity while allowing a high degree of freedom to the AM.

\subsection{Global optimization}
\label{sec:atmoswing-optimization}

The sequential calibration cannot optimize all parameters of the AM, as several are manually selected, such as the meteorological variable, the pressure level and the temporal windows. Testing multiple combinations of these can be cumbersome, especially when considering multiple predictors within the same level of analogy. Moreover, the parameters are calibrated successively within the same level of analogy, and in-between them.

We thus decided to explore optimization techniques that can handle all parameters fully automatically. We first assessed the \citet{Nelder1965a} method, which was found to be unsuitable for this purpose. Due to the complexity of the AM and its non-linearity, a technique of global optimization is required. We thus implemented Genetic algorithms in AtmoSwing Optimizer in order to perform a global optimization of the AM. This article is not about discussing the results of such a technique, but describing the details of the implementation. The demonstration of the benefit brought by such an approach is the topic of another article. The details of the implementation being numerous, the description is provided in the next dedicated section (\ref{sec:gas}).

\section{Genetic algorithms}
\label{sec:gas}

Genetic algorithms (GAs) come from the world of stochastic optimization, more specifically from metaheuristic approaches. These are stochastic iterative algorithms that behave like search algorithms by exploiting the characteristics of a problem and are particularly suitable for complex parameter spaces.

Genetic algorithms are part of the family of evolutionary algorithms \citep{Back1993b, Schwefel1993}, which get inspiration from some mechanisms of biological evolution, such as reproduction, genetic mutations, chromosomal crossovers, and natural selection. GAs are the most used technique from evolutionary algorithms \citep{Back1993b}, and they are constantly improving \citep{Haupt2004}. However, with time, the different methods of evolutionary algorithms tend to be similar and share many commonalities \citep{Back1996b, Haupt2004}.

The method was originally developed by \citet{Holland1992b} and popularized by \citet{Goldberg1989}. Unlike a linear or local optimization, GAs seek the global optimum on a complex surface, theoretically without restriction, but with no guarantee to reach it.

\subsection{Basic concepts of the genetic algorithms}

GAs mimic the evolution of a population of individuals in a new environment, by applying rules based on natural processes, such as DNA mutation, chromosomes crossover, natural selection, etc. It simulates the fact that in a natural environment, the most suitable individuals tend to survive longer, to reproduce more easily, and so to influence coming generations by providing genes that provide some good performance in a certain domain. Generation after generation, the DNA mixes and the strong genes cumulate in some individuals \citep{Beasley1996a}. Globally, the fitness of the population to its environment increases, while retaining enough variety to not converge too quickly to a local optimum.

Applications of GAs are very diversified as they can handle many parameters of various types \citep{Joines1996a}, even with very complex cost surfaces \citep{Haupt2004}. GAs work remarkably well with intervariable dependences \citep{Haupt2004}. The objective function to optimize (often named fitness function in this context) can be of different types (mathematical function, experimental or numerical modeling). Only the resulting value is used for optimization. Indeed, these algorithms do not require any knowledge of the problem, which can be used as black-box, but they must be adapted in order to perform optimally.

By means of the reproduction operator and the natural selection, the GAs focus on the most promising regions of the parameter space \citep{Holland1992b}. Points (parameter sets) are densified in these areas because the strong genes of the best individuals propagate from generation to generation.

Two conditions guarantee in theory the convergence to the global optimum \citep{Zitzler2004a}: (1) Parameters mutations that can allow to explore the entire parameter space, thereby ensures that any value can be achieved with a non-zero probability. (2) A rule of elitism ensuring that an optimal solution cannot be lost or damaged.

\subsection{Structure and operators}

The GAs optimize a population of individuals (parameter sets). Each individual contains a chromosome (parameters of the AM in our case). We call gene every parameter that constitutes the chromosome. The parameters to optimize have long been coded in binary form and assembled as strings in the canonical GAs \citep{Goldberg1989}. Encoding and decoding steps were needed to transform the variable from its floating-point representation into its binary representation, and vice versa, which introduce quantification errors \citep{Haupt2004}. According to \citet{Holland1992b}, working with binary chromosomes was supposed to be more efficient \citep{Goldberg1990a, Back1993b}. However, more and more applications use floating-point representations, allowing to avoid the coding and decoding steps and the quantification errors \citep{Haupt2004}, and which also often resulted in a performance improvement \citep{Goldberg1990a}. It is thus now considered that for continuous variables, a floating-point representation is more suited \citep{Michalewicz1996, Herrera1998a, Haupt2004, Back1996b, Gaffney2010a}.

There are numerous implementation variants of GAs often optimal for a given problem \citep{Hart1991a,Schraudolph1992a}. However, the structure of the method (Figure \ref{fig:structure_gas}) resulting from the work of \citet{Holland1992b} is common to most applications \citep{Back1993b}. The divergences are the operators implementation, through significantly different algorithms, which has an important effect on the results \citep{Gaffney2010a}.

All operators we used and their options, applied to real coding, are described in the following sections. Many other operators exist, but we will only present the ones we evaluated.

\subsubsection{Genesis of the population}

The first step of the optimization is to generate the initial population. A population is a set of $N$ individuals (each of which represents a point in the space of potential solutions, a parameter set of the AM in our application) that we are going to make evolve. A generation is the population at a given time.

A random initialization based on a uniform distribution is the most current version. The size $N$ of the population is often a compromise between the computation time and the quality of the solution. $N$ must allow sufficient sampling of the solutions field \citep{Beasley1996a}, and should thus vary as a function of chromosome size (ie the number of parameters to be optimized).

\subsubsection{Natural selection}

Natural selection is performed on the basis of the values of the objective function. The selection allows to only keep a certain part of the population, usually half ($N/2$), which can access the mating pool (intermediate generation with $N_{mp}$ members). If $N_{mp}$ is too high, the reproduction rate is too low, whereas if it is too small, the strong traits of individuals do not have the ability to accumulate in the same chromosome \citep{Haupt2004}. Several techniques exist, such as:

\begin{itemize}
	\item \textbf{$N_{mp}$-elitism} \citep{Michalewicz1996}: the population is sorted according to the value of the objective function and only the better half is preserved.

	\item \textbf{Tournament selection} \citep{Michalewicz1996, Zitzler2004a}: two individuals are randomly selected and fight. The one with the highest score is chosen, but with a certain probability, in order to reduce the selection pressure. This procedure is repeated until the mating pool is full. Individuals can be selected several times, and thus be represented several times in the mating pool.
\end{itemize}

\subsubsection{Selection of the couples}

Individuals of the mating pool can reproduce. It begins with the selection of pairs (the parents). The techniques implemented in this work are the following:

\begin{itemize}
	\item \textbf{Rank pairing}: individuals are gathered in pairs according to their rank (classified on the performance scores). Consecutive ranks are put together (odd rows are associated with even rows). This approach is easy to achieve, but does not look like a natural process.

	\item \textbf{Random pairing}: two individuals are randomly selected to form a couple, according to a uniform law.

	\item \textbf{Roulette wheel weighting}: the roulette technique refers to gambling. But unlike casino roulette, this one is biased. Each individual is associated with a sector of the wheel with a certain opening angle, which is its probability of selection \citep{Haupt2004}. The probability assigned to the individuals is proportional to their fitness (objective function), so that the most adapted individuals have the greatest probability of reproduction. There are two techniques for weighting the individuals of the mating pool:

	\textit{Roulette wheel weighting on rank}: the probability of each individual depends on its rank $n$:
	\begin{equation}
	p_{n}=\dfrac{N_{mp}-n+1}{\sum^{N_{mp}}_{n=1}n}
	\label{equation_mating_rank_weighting}
	\end{equation}

	\textit{Roulette wheel weighting on fitness}: the selection probability is calculated based on the value of the objective function. This approach gives more weight to the best individuals when the distribution of scores is wide, while the weight is almost similar when all individuals have approximately the same score \citep{Haupt2004}. The probability $p_{n}$ of each individual is calculated by the equation \ref{equation_mating_score_weighting}:
	\begin{equation}
	p_{n}=\frac{score_{n}-score_{N_{mp}}}{\sum_{n=1}^{N_{mp}} (score_{n}-score_{N_{mp}})}
	\label{equation_mating_score_weighting}
	\end{equation}
	In our application, the last individual ($N_{mp}$) has zero probability of being selected.

	\item \textbf{Tournament selection}: This operator is similar to the one used in natural selection, but is applied here for the successive selection of each parent. To select a parent, a number of individuals (2 or 3) are randomly picked and the best is kept. This operation is performed twice, once for each partner. This approach imitates the breeding competition in nature \citep{Haupt2004}.
\end{itemize}

\subsubsection{Chromosome crossover}

Once the two parents are selected for breeding, they combine their chromosomes and produce two children, bringing the number of individuals in the population back to $N$ (the parents also return back in the total population in order to complement the next generation). The combination of chromosomes is carried out using a crossover operator, thereby generating two offspring having characteristics derived from both parents. Chromosome crossover widens the search space and favours the combination of strong genes, which can result in more suited children. It allows a mixing of genes and accumulation of positive mutations.

The evaluated crossover operators are the following:

\begin{itemize}
	\item \textbf{Single-point crossover}: a crossover point is randomly chosen for the pair. The genes (our parameters) located after that point are exchanged in between the two chromosomes.

	\item \textbf{Two-point crossover}: works like the single-point crossover, but there are two intersections defining the segments to exchange. This approach, which significantly extends the search space for the children, is considered more efficient than the previous \citep{Beasley1993a, Haupt2004}.

	\item \textbf{Multiple-point crossover} \citep{DeJong1975a}: it is a generalization of the previous, with a number of crossover points up to the number of genes.

	\item \textbf{Uniform crossover} \citep{Syswerda1989}: for each gene of the chromosome, it is randomly chosen to exchange or not the values between the parents.

	\item \textbf{Binary-like crossover} \citep{Haupt2004}: chromosome crossover on a binary coding can generate new values for variables located at intersection points, since the crossovers are applied at the bit level, thus often within a gene. This is not the case for the floating-point representation, since the crossover is performed between the genes. To reproduce the behaviour present in the original algorithms, which introduces new information, \citet{Haupt2004} propose an operator that combines standard crossover with an interpolation approach. The genes located after a crossover point are exchanged, but the gene located at the intersection is modified as follows (equation \ref{equation_mating_as_binary}):
	\begin{equation}
	\left\lbrace \begin{array}{l} 
	g_{o1,n} = g_{p1,n} - \beta (g_{p1,n} - g_{p2,n}) \\
	g_{o2,n} = g_{p2,n} + \beta (g_{p1,n} - g_{p2,n}) \\
	\end{array} \right.
	\label{equation_mating_as_binary}
	\end{equation}
	where $g_{o1,n}$ and $g_{o2,n}$ are the $n$-th gene of the two new offspring, and $g_{p1,n}$ and $g_{p2,n}$ are those of the two parents. $\beta$ is a random value between 0 and 1.

	\item \textbf{Blending method} \citep{Radcliffe1991a}: in this approach, instead of exchanging the genes in between the chromosomes after one or multiple crossover points, these are combined by linear combination (equation \ref{equation_mating_blending_method}). The genes of the parents are blended together using a random value ($\beta$) that can be unique for the whole chromosome, or that can change for every gene. The genes of the offspring are bounded by the genes of the parents, no value can be out of their range.
	\begin{equation}
	\left\lbrace \begin{array}{l} 
	g_{o1,n} = \beta g_{p1,n} + (1-\beta)g_{p2,n} \\ 
	g_{o2,n} = (1-\beta) g_{p1,n} + \beta g_{p2,n} \\
	\end{array} \right.
	\label{equation_mating_blending_method}
	\end{equation}

	\item \textbf{Linear crossover} \citep{Wright1991a}: in order to allow the genes to take values outside the interval defined by the parents, a method of extrapolation is necessary. Linear crossover introduces such an approach, and produces three children from two parents, following equation \ref{equation_mating_linear_crossover}. Less couples are required in order to fill up the generation.
	\begin{equation}
	\left\lbrace \begin{array}{l} 
	g_{o1,n} = 0.5 g_{p1,n} + 0.5 g_{p2,n} \\ 
	g_{o2,n} = 1.5 g_{p1,n} - 0.5 g_{p2,n} \\ 
	g_{e3,n} = - 0.5 g_{p1,n} + 1.5 g_{p2,n} \\ 
	\end{array} \right.
	\label{equation_mating_linear_crossover}
	\end{equation}

	\item \textbf{Heuristic crossover} \citep{Michalewicz1996}: it is a variation of the latter methods that relies on the following equation:
	\begin{equation}
	\left\lbrace \begin{array}{l} 
	g_{o1,n} = \beta (g_{p1,n} - g_{p2,n}) + g_{p1,n} \\
	g_{o2,n} = \beta (g_{p2,n} - g_{p1,n}) + g_{p2,n} \\
	\end{array} \right.
	\label{equation_mating_heuristic_crossover}
	\end{equation}

	\item \textbf{Linear interpolation}: unlike previous techniques, this technique does not rely on crossover points, but on a linear interpolation on every gene of the couple (equation \ref{equation_mating_linear_interpolation}).
	\begin{equation}
	\left\lbrace \begin{array}{l} 
	c_{o1} = c_{p1} - \beta (c_{p1} - c_{p2}) \\
	c_{o2} = c_{p2} + \beta (c_{p1} - c_{p2}) \\
	\end{array} \right.
	\label{equation_mating_linear_interpolation}
	\end{equation}
	where $c_{o1}$ and $c_{o2}$ are the full chromosomes of the offspring, and $c_{p1}$ an $c_{p2}$ are the ones of the parents. As before, $\beta$ is a random value between 0 and 1, and is here the same for every gene.

	\item \textbf{Free interpolation}: this technique performs interpolation on each gene, like the previous one; but in this case, the weighting factor changes for each gene:
	\begin{equation}
	\left\lbrace \begin{array}{l} 
	c_{o1} = c_{p1} - [\beta_{1} (g_{p1,1} - g_{p2,1}), \beta_{2} (g_{p1,2}\\
	~~~~~~~~~~~~ - g_{p2,2}), ..., \beta_{Ng} (g_{p1,N_{g}} - g_{p2,N_{g}})] \\
	c_{o2} = c_{p2} + [\beta_{1} (g_{p1,1} - g_{p2,1}), \beta_{2} (g_{p1,2}\\
	~~~~~~~~~~~~ - g_{p2,2}), ..., \beta_{Ng} (g_{p1,N_{g}} - g_{p2,N_{g}})] \\
	\end{array} \right.
	\label{equation_mating_free_interpolation}
	\end{equation}
	where $N_{g}$ is the number of genes, and $\beta$ is here independent between the genes.

\end{itemize}

Many other methods or variations exist, combining the advantages of different approaches. The performance of the variants being related to the problem to be addressed, we can not identify a priori the best technique for our application.

\subsubsection{Mutation}

The combination of strong genes by the operator of chromosomes crossover is theoretically the most important operating mechanism in the conventional GAs \citep{Holland1992b,Back1993b}. However, many studies identify the mutation process as main operator, and crossovers as secondary \citep[see][]{Back1992a,Back1996a,Back1996b,Smith1997a,Deb1999,Haupt2004,Costa2005a,Costa2007a}.

The mutation operator is a direct modification of genes. In a binary coding, it is implemented as an inversion of some bits in a chromosome, while in real coding, it is done by changing the gene values. Mutations add diversity to the population and prevent a freeze of the evolution, or a genetic drift to a local optimum. Thus, it makes the convergence to the global optimum theoretically possible \citep{Beasley1993a}, as they allow exploring beyond the current region of the parameter space. They therefore help preventing the algorithm to converge too quickly to a local optimum and bring new characteristics that were not present in the original population \citep{Haupt2004}.

The evaluated and developed mutation operators are the following:

\begin{itemize}
	\item \textbf{Uniform mutation}: The mutation rate is constant and equal for every gene of each individual; they all have the same probability to mutate. When a gene is selected for mutation, a new random value is assigned, according to a uniform law.

	\item \textbf{Variable uniform mutation} \citep{Fogarty1989}: a variable mutation rate over the generations was first suggested by \citet{Holland1992b} and evaluated by \citet{Fogarty1989}. It improved significantly the performance of GAs. In most applications, the mutation rate decreases with the generations, in a deterministic and global (for all individuals) manner \citep{Back1992b}. Its optimum configuration depends on the size of the chromosomes, of the properties of the objective function, and of the population size \citep{Back1992b}. We implemented this operator according to equation \ref{equation_mutation_uniformvariable}.
	\begin{equation}
	p_{n,G} = p_{G_{0}}+\left( \dfrac{p_{G_{0}}-p_{G_{m,p}}}{G_{m,p}} \right) min\left\lbrace G,G_{m,p}\right\rbrace 
	\label{equation_mutation_uniformvariable}
	\end{equation}
	where $p_{n,G}$ is the mutation rate (probability) of the gene $n$ for generation number $G$, $G_{m,p}$ is the maximum number of generations during which the mutation rate varies. $p_{G_{0}}$ is the initial mutation probability, and $p_{G_{m,p}}$ is the final one. $p_{G_{0}}$, $p_{G_{m,p}}$ and $G_{m,p}$ are the three controlling parameters of the operator. The evolution of the mutation rate is linear.

	\item \textbf{Constant normal mutation}: many use normal distributions to generate new values. The gene $g$ that mutate becomes:
	\begin{equation}
	g' = N(g,\sigma^{2})
	\label{equation_mutating_normal_distribution}
	\end{equation}
	where $\sigma$ is the standard deviation of the distribution. The disadvantage of this technique is that an accurate value of $\sigma$ must be chosen \citep{Haupt2004}, which is impossible to know beforehand.

	\item \textbf{Variable normal mutation}: with the same logic that the variable uniform mutation, we tested a mutation operator using a normal distribution with a variable mutation rate and standard deviation. The mutation rate is calculated with equation \ref{equation_mutation_uniformvariable}. On the same principle, we decrease linearly the standard deviation over the generations:
	\begin{equation}
	\sigma_{n,G} = \sigma_{G_{0}}+\left( \dfrac{\sigma_{G_{0}}-\sigma_{G_{m,\sigma}}}{G_{m,\sigma}} \right) min\left\lbrace G,G_{m,\sigma}\right\rbrace 
	\label{equation_mutation_normalvariable}
	\end{equation}
	where $\sigma_{n,G}$ is the standard deviation of gene $n$ et generation number $G$, $\sigma_{G_{0}}$ is the initial standard deviation, $\sigma_{G_{m,\sigma}}$ is the final standard deviation, $G_{m,\sigma}$ is the maximum number of generations during which the standard deviation varies. $p_{G_{0}}$, $p_{G_{m,p}}$, $G_{m,p}$, $\sigma_{G_{0}}$, $\sigma_{G_{m,\sigma}}$ and $G_{m,\sigma}$ are the six parameters of the method.

	\item \textbf{Non-uniform mutation} \citep{Michalewicz1996}: two random numbers are picked based on a uniform law: $r_{1}$, which determines the direction of the change, and $r_{2}$, which determines its magnitude. The new value of the gene is given by the following equation:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} \left(1 - \dfrac{G}{G_{m}} \right)^{2} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} \left(1 - \dfrac{G}{G_{m}} \right)^{2} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{equation_mutation_nonuniform_original}
	\end{equation}
	where $a_{n}$ is the is the lower bound of the $n$-th gene, $b_{n}$ its upper bound, $G$ the present generation, and $G_{m}$ the maximum number of generations.

	We adapted this operator for our application, which is not based on a predefined number of generations:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} \left(1 - \min \left\lbrace \dfrac{G}{G_{m,r}}, 1 \right\rbrace \left(1-\omega\right) \right)^{2} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} \left(1 - \min \left\lbrace \dfrac{G}{G_{m,r}}, 1 \right\rbrace \left(1-\omega\right) \right)^{2} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{equation_mutation_nonuniform}
	\end{equation}
	where $G_{m,r}$ is the maximum number of generations during which the magnitude of the research varies, and $\omega$ is a threshold chosen by the user to maintain a minimum search radius when $G>G_{m,r}$. During the first generations, the exploration extent covers the entire parameter space. However, this area is reduced over generations, allowing exploitation of local solutions.

	\item \textbf{Individual adaptive mutation rate} \citep{Back1992a}: based on the ideas of Evolution Strategies \citep[see][]{Rechenberg1973, Schwefel1981}, \citet{Back1992a} introduced a concept of self-adaptive genetic algorithms. The idea is to distribute control parameters within individuals themselves, which partially decentralize control of the evolution. It allows reducing the parametrization of GAs and introducing a notion of self-management. The first approach is the introduction of a mutation rate per individual, that mutates itself under its own probability \citep{Back1992a}. Then, the eventual new rate is used to mutate the genes of the individual. Thus, as this rate decreases, it will have less probability of being itself mutated. This approach is close to the natural adaptation phenomena. A population less suited to its environment is changing faster than better adapted populations. Mutations are performed according to a constant uniform distribution. The initial mutation rates are randomly chosen \citep{Back1992a} and the method has no parameter. Other approaches exist to introduce a self-adaptation \citep[see][]{Smith1997a,Deb1999,Deb2001a}.

	\item \textbf{Individual adaptive search radius}: based on the ideas of the non-uniform mutation, we introduce a search radius in the approach of individual adaptive mutation rates. This search radius $r_{a}$, bounded between 0 and 1, is also adaptive and behaves similarly to the adaptive mutation rates. In order to separate its evolution from the one of the mutation rate, its own value is considered initially as a self-mutation rate to eventually mutate before being used as a normalized search radius. The value of a mutated gene is given by the following equation, which is a simplification of the non-uniform mutation:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} r_{a} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} r_{a} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{equation_mutation_rayon_adaptatif}
	\end{equation}
	where $r_{1}$ and $r_{2}$ are randomly selected, in the same way as for the non-uniform mutation. No external parameter is therefore necessary.

	\item \textbf{Chromosome of adaptive mutation rate} \citep[\textit{n adaptative mutation rate},][]{Back1992a}: analogously to the individual adaptive mutation rate, this approach leaves the control of the evolution rate to the individuals themselves. The difference here is that each gene has a specific mutation rate. The main advantage is that the tuning of the mutation can be much more precise \citep{Smith1997a}. We therefore consider a second chromosome containing the mutation rate for each gene of the first chromosome. The operations of mutation and self-mutation are similar to the case of the individual adaptive mutation rate, but in a distributed way, within the chromosome. Another difference is that we apply the same crossover operations as those applied to the first chromosome, and this for the same crossing points. Thus, during an exchange of genes, children also inherit the mutation rates specific for each of these genes.

	\item \textbf{Chromosome of adaptive search radius}: we introduced this operator that combines the operations of the chromosome of adaptive mutation rate to our adaptive search radius approach. Similarly, an individual has 3 chromosomes: the first containing the values to be optimized, the second containing the distributed mutation rate, and the last one, the distributed search radius. Again, no external parameters are required.

	\item \textbf{Multi-scale mutation}: finally, we developed another approach, that is also based on the search radius concept. However, the latter is not decreasing with time. Methods based on a reduction of the mutation rate or radius simulate a transition from the exploration phase to the exploitation one. The idea is consistent as long as we are confident that the algorithm will converge towards the global optimum. Indeed, once the algorithm is in the exploitation mode, it is very unlikely to go out of the region it converges to. We wanted to test an approach that combines both exploration and exploitation during the whole optimization. Thus, we considered the search radius $r_{a}$ of equation \ref{equation_mutation_rayon_adaptatif} as a random value for each individual, but restricted to 4 equiprobable values: 1, 0.5, 0.1, 0.02. The only external parameter is the mutation rate which is fixed.

\end{itemize}

When the gene to mutate is represented by a list of distinct values (eg meteorological variable or analogy criterion), the random choice of a new value is always based on a uniform distribution, without notion of search radius. There is indeed no meaning to use operators based on principles of proximity when the latter does not exist.

\subsubsection{Elitism}

We used a process of elitism on the natural selection as well as on mutations. This ensures the survival of the best individual so that we do not lose a better solution. This approach is very common in the field of GAs \citep{Haupt2004}. After the natural selection operator, if the best individual has not been selected, it is copied to the mating pool instead of an individual randomly picked. After mutation, if the best individual has mutated and if its new version has a lower score than the original, the latter is reinserted instead of an individual randomly chosen.

\subsubsection{Ending the optimization}

The convergence check determines whether the solution is acceptable and if the algorithm may stop. The stopping criteria are not often well documented in GAs case studies. We chose to stop the optimization if the best individual does not change for $x$ generations. This value should not be too low to allow the algorithm to escape from a local optima. In addition, the rate of improvement decreases with the progression of the optimization. It is thus common that the best individual does not evolve over several generations when we get closer to the global solution. We chose a value of $x=20$ generations.

\subsection{Implementation and constraints}

Some constraints need to be taken into account. For example, when a crossover or a mutation operation results in a parameter value standing out of the authorized bounds, it has to be brought back within the limits. Moreover, the parameters are of different nature: some are continuous, such as the weight, some are discretized, such as the analogues number, or the spatial windows, and finally, some are independent elements in an array, such as the selection of the meteorological variable. New values resulting from the optimizer need to respect the type of data it represents.

Other constraints exist in between the parameters, such as the temporal window of the moisture index that has to be consistent in between the relative humidity and the precipitable water. Another example is the weighting of the different pressure levels which has to be normalized.

GAs are very computationally intensive because they require many evaluations of the objective function. These assessments are very long in our application, as they require calculating and assessing a forecast for every day of the calibration period. In order to reduce the computation time, we avoid recalculating the score of an individual who has previously been evaluated and that has not changed. We keep the score of each individual living in the selection until it mutates.

The assessment (calculation of the objective function) of each member of the population of a generation is completely independent and can be performed in parallel on different processors of a computer \citep{Alliot2005}. We implemented this technique and the resulting time savings was very important. In order to perform optimizations for multiple time series, the use of a cluster is a necessity, which our code allows.

\subsection{Recommendations of parametrization}

The GAs parametrization, such as the mutation rate, population size, natural selection options, and so on, is difficult given the high number of existing variants, each developed for a specific problem \citep{Haupt2004, Costa2007a}. This parametrization depends on the objective function, implementation variants, the range of the parameters to be optimized, and performance indicators. Thus, different studies suggest very different parametrization.

A key element of the parametrization of GAs is finding the right balance between exploration and exploitation \citep{Back1992a, Smith1997a}. Exploration is characterized by a relatively high probability to assess the regions of the parameter space that have not yet been visited. This probability must be sufficiently large at the beginning of the optimization, so that the algorithm is capable of identifying the region where the global optimum is located. Exploitation is characterized by a local search in an area of interest, and generally makes small movements. The latter is interesting to refine the results at the end of the optimization.

\citet{DeJong1975a} and \citet{Grefenstette1986} compared different implementations and parametrizations of GAs on functions of varying complexity. They observed that a small population size improves the initial performance, while a large population improves long-term performance. They also observed that the ratio of the population to keep for the mating pool is around 50\% (45\% to 60\%).

Values of the mutation rate varies broadly between the studies: from 0.001 \citep{DeJong1975a} to 0.2 \citep{Haupt2004}. \citet{Back1996b} showed that mutation rates higher than the usual ranges are more optimal at the beginning of the optimization, allowing further exploration. The combination of a small population and a high mutation rate works best for the first generations \citep{DeJong1975a, Back1996b, Haupt2004}, but as we could observe, it does not guarantee the quality of the final result. Incremental approaches with varying mutation rates are certainly more optimal but more complex to implement \citep{Back1996a, Back1996b}.

\subsubsection{Comparison process and results}

One of our goals being to make recommendations of parametrization in order to optimize the AM, we proceeded systematically. The results are summarized hereafter \citep[see][for the details]{Horton2012a}. We used concepts from the factorial design approach \citep[see eg.][]{Costa2005a,Costa2007a,Mariano2010a}, which is sometimes used for comparative analysis of different parametrizations of GAs. We processed by stages, analysing in details and in a systematic way every variants of the implemented operators, in combination with multiple other options and parameters in order to take into account eventual co-dependencies.

In order to evaluate a combination of operators/options, we processed 10 optimizations for one parametrization of GAs. The performances were characterized by four indicators:

\begin{itemize}
	\item mean score: average of the final scores of the 10 optimizations,
	\item convergence: the number of optimizations that converged to a supposed global optimum,
	\item number of generations: characterization of the convergence speed,
	\item number of evaluations of the objective function: characterization of the required calculation time (more realistic than the number of generations).
\end{itemize}

This comparison required tens of thousands of optimizations that were performed on a cluster of the University of Lausanne. The results, detailed in \citet{Horton2012a} are synthesized hereafter:

\begin{itemize}
	\item \textbf{Population size}: we found the following ranges to be accurate in average:

	$50<N<100$ for a very simple implementation of the AM (1 level of analogy with 2 pressure levels),

	$N\approx200$ for a bit more complex AM (1 level of analogy with 4 pressure levels, or 2 level of analogy with less pressure levels),

	$N\approx500$ for significantly more complex AMS (2-3 levels of analogy with 4 pressure levels for the atmospheric circulation, and 2 to 4 levels for the moisture analogy),

	We didn't find any improvement with $N>500$, the results were even surprisingly of a lower quality. However, this cannot be generalized and depends on the AM to optimize, and supposedly on the characteristics of the processes generating the precipitations in a given region.

	\item \textbf{Natural selection}: this operator has no significant influence, and both tested implementations work fine, with a slightly better performance for the ratio-elitism.

	\item \textbf{Selection of couples}: 6 variants of the couples selection were assessed. The performance of these variants are relatively close, both in terms of score, convergence, and number of evaluations. The random pairing performed the most poorly, when the tournament selection with 3 candidates was slightly superior. The roulette wheel weighting is not far behind, but it is less effective in terms of convergence and number of evaluations. This operator has not a significant role in our application.

	\item \textbf{Chromosomes crossover}: we compared 21 different options of the crossover operators. This analysis revealed some slightly better options, some bad ones, and many averages. Among the bad operators, we find first the heuristic crossover, which is also more demanding in number of evaluations, as well as the linear crossover. Binary-like crossovers (especially with 2 points of intersection, whether $\beta$ is shared or not) are significantly better than the others, especially in terms of convergence. The two points crossover is relatively close. Other operators can be considered usable, yet may not be optimal. Once again, this operator is not the key of the GAs parametrization.

	\item \textbf{Mutation}: we compared the 10 mutation operators with different options, bringing the number of variations of this operator up to 110. We immediately observed that the mutation operator has a very important role on the performance of the optimizations of the AM, and that the other reproduction operators seem of secondary importance. This observation is in line with the work of \citet{Back1996a}, who argues for the importance of mutation over reproduction. He even suggests, in opposition with the theory of \citet{Holland1992a}, that chromosomes crossovers have mostly a corrective role of mutation operations. Various studies have also identified the importance of the mutation operator relatively to reproduction \citep[see eg.][]{Back1992a, Back1996b, Smith1997a, Deb1999, Haupt2004, Costa2005a, Costa2007a}.

	The mutation operators based on a variable normal or uniform law work very poorly and are difficult to configure. We then observe many operators more or less with the same scores and requiring a variable amount of assessments. The convergence analysis allows us to highlight three best operators:

	\textit{Non-uniform mutation} \citep{Michalewicz1996}: this operator is good in terms of convergence, mainly when the number of parameters to optimize is rather low. The number of required evaluations, however, can be quite substantial. The main disadvantage of the non-uniform mutation is the complexity of its parametrization, which is difficult to estimate a priori. These parameters must be carefully chosen to be in line with the evolution rate of the population, and are therefore dependent on the problem being treated. We could observe that the $\omega$ coefficient does not influence performance. The role of $G_{m}$ is rather difficult to judge, but does not seem essential. The mutation rate was found to be important. The difficulty is that the optimal value seems to be very case-related. Indeed, by even changing the precipitation time series (ie optimizing for another subregion), but not the complexity of the AM, the optimal mutation rate changes, making it impossible to estimate in advance.

	\textit{Chromosome of adaptive search radius}: unlike the previous one, our new operator is very robust, as it requires no option and is auto-adapting. It may be sometimes a little bit slower for simple problems, but does not require parametrization, which is an important advantage. It is interesting to notice that our insertion of an extra chromosome representing the search radius gives better performance than other self-adaptive operators (such as, for example, the chromosome of adaptive mutation rate).

	\textit{Multi-scale mutation}: finally, our multi-scale mutation, which also performs pretty well, can as well be seen as fairly robust, since it requires only one parameter, the mutation rate. However, it can also be difficult to estimate a correct value a priori.

	It may be wise to perform multiple optimizations and to consider these three operators in parallel in order to obtain results from algorithms that are either sometimes more efficient or more robust. It is interesting to note that the three best techniques incorporate a notion of search distance. It is likely that this notion is the key to these algorithms, for our application, and allows them to initially explore the parameter domain, and then to converge. The search radius in fact directly represents the notion of transition between exploration and exploitation, in our opinion more than a possible evolution of mutation rates.

	\item \textbf{Other options}: A ratio of 50\% for the mating pool seems to be a good choice.

\end{itemize}

\subsubsection{Recommended parametrization of GAs}

We evaluated the optimization by genetic algorithms on AMs of varying complexity, with a large number of combinations of operators to be able to make recommendations for optimizing the AM. Our conclusions are:

\begin{itemize}
	\item The optimization does not systematically converge to the global optimum  (but still often nearby), which is why we recommend doing several optimizations in parallel in order to compare the results, analyse the convergence, and keep the best.

	\item The population size should be in accordance with the complexity of the AM to optimize: from 50 for the simple ones, up to 500 for the most complex AMs.

	\item The value of the ratio for the intermediate population is not so important, and value of 50\% seems quite appropriate.

	\item Ratio-elitism is slightly better than tournaments for the natural selection operator, but it is not decisive.

	\item The performance of the operators for the couples selection perform relatively similarly. The roulette wheel weighting and the tournament selection are more efficient in terms of convergence and required number of evaluations.

	\item Most crossover operators have relatively similar performance. Binary-like crossover with two points of intersection are better than others, especially for convergence.

	\item Mutation has a clearly dominant influence. Three mutation operators stand out, two of which we have developed: the non-uniform mutation, the multi-scale mutation, and the chromosome of adaptive search radius. The latter is the most robust as it has no controlling parameter.

\end{itemize}

In order to make an optimization of the AM with genetic algorithms, it may be wise to consider these three mutation operators in parallel. We would then combine algorithms that are sometimes faster to other that are more robust. In order to be confident in the optimized AMs, we propose using a set of the following mutation operators:

\begin{itemize}
	\setlength\itemsep{-4px}
	\item 1x non-uniform, $p_{mut}=0.1$, $G_{m}=50$, $\omega=0.1$
	\item 1x non-uniform, $p_{mut}=0.1$, $G_{m}=100$, $\omega=0.1$
	\item 1x non-uniform, $p_{mut}=0.2$, $G_{m}=100$, $\omega=0.1$
	\item 1x multi-scale,  $p_{mut}=0.1$
	\item 2x chromosome of adaptive search radius
\end{itemize}

\section{Perspectives}

%\citet{Thevenot2004} a étudié plusieurs aspects de la méthode des analogues en utilisation opérationnelle. Le premier a consisté à intégrer les prévisions d'ensemble du Centre Européen de Prévision Météorologique à Moyen Terme (CEPMMT, ou \textit{European Centre for Medium-Range Weather Forecasts}, ECMWF). Cette approche permet de tenir compte des incertitudes sur les prédicteurs issus du modèle global. Son implémentation est une combinaison des journées analogues retenues, associées à chacune des traces. La distribution des précipitations est ensuite établie sur l'ensemble des analogues. Ainsi, certaines journées peuvent être représentées plusieurs fois. À partir d'une échéance de 4 jours, la prévision sur l'ensemble des traces se montre plus performante que celle effectuée sur le seul contrôle déterministe.

%Il a aussi effectué une calibration du nombre d'analogues à retenir en fonction de l'échéance pour la prévision déterministe, ainsi que pour la prévision d'ensemble. Pour la prévision déterministe, le nombre d'analogues pour les deux premières échéances est identique à celui de la prévision parfaite \citep[calibration sur les réanalyses,][]{Klein1963}, puis croît de manière importante jusqu'à tendre vers l'ensemble de la climatologie. Cette optimisation a été effectuée sur la méthode à un niveau d'analogie (analogie de circulation), mais n'a pas pu être entreprise de manière aussi rigoureuse sur le second niveau d'analogie.

%\citet{Marty2010} a appliqué plusieurs scores de validation aux quantiles de la distribution des prévisions opérationnelles de la méthode des analogues. Il en a conclu que la détection de l'occurrence des pluies était mieux prévue par les quantiles médians (50 à 70~\%), avec un optimum à 60~\%. Toutefois, pour les pluies significatives, le quantile 90~\% donnait les meilleures performances. En comparant les résultats des analogues et des prévisions d'ensemble, pour le quantile 90~\%, il a observé que, à partir de $P = 0.16 \cdot P10$, les analogues donnaient trop d'alertes, alors que les ensembles n'en émettaient pas assez. La prévision opérationnelle par analogie était meilleure que celle des ensembles considérés, et ceci particulièrement pour les forts cumuls \citep{Marty2010}. Le gain des analogues provient principalement d'une meilleure justesse, pour une moins bonne finesse. Au contraire, les prévisions d'ensemble sont très fines, mais manquent un peu de justesse.

%Lors de l'analyse du biais des résultats de la méthode des analogues en opérationnel (également des ensembles), il a observé que la fréquence prévue des pluies nulles ($F(0)$) était biaisée et sous-estimait la fréquence réelle des précipitations nulles. Au contraire, les fréquences prévues des pluies positives ($F_{+}$) surestimaient les fréquences observées. Ainsi, par exemple, le quantile 50~\% de la distribution prévue correspondrait au quantile réel 32~\%, et le 90~\% prévu au 81~\% réel. Il a donc proposé une correction des biais.

%En analysant le biais de la méthode des analogues en prévision parfaite (voir section \ref{sec:etat_art:calibration}), il a observé que $F(0)$ et $F_{+}$ n'étaient que très légèrement biaisés, et que les biais observés précédemment avaient donc une origine externe (l'échantillon des prévisions du modèle numérique global utilisé). Toutefois, en appliquant un conditionnement par $F(0)$, il est apparu que \og \textit{si la journée cible est prévue pour être plutôt humide ($F(0) \leq 0.10$), les fréquences de $F_{+}$ sont surestimées (et les quantiles de précipitations sont dès lors sous-estimés)}\fg{}, et au contraire, que \og \textit{si elle est prévue pour être plutôt sèche avec une probabilité moyenne ($F(0) > 0.20$), alors les fréquences $F_{+}$ sont sous-estimées (et les quantiles de précipitations sont dès lors surestimés)}\fg{}. Il a donc proposé une méthode de correction a priori en appliquant un coefficient correctif sur chaque fréquence quantilique en fonction de $F(0)$, puis en déterminant la distribution a posteriori, ajustée selon la loi Gamma à deux paramètres. Cette correction s'est avérer améliorer les performances des prévisions pour les journées à fortes précipitations.

%Il a encore proposé une désagrégation temporelle à 6 et 12~h des prévisions de la méthode des analogues pour une utilisation opérationnelle dans une chaîne hydrométéorologique. La meilleure méthode identifiée consistait à utiliser les cumuls prévus par la méthode des analogues à un pas de temps journalier, et de leur appliquer la chronologie des prévisions d'ensemble, lesquels reproduisent bien la répartition infrajournalière.

\conclusions  %% \conclusions[modified heading if necessary]
% TODO: write

\appendix
%\section{}    %% Appendix A

%\subsection{}                               %% Appendix A1, A2, etc.

\begin{acknowledgements}
Thanks to Lucien Schreiber and Richard Metzger for their programming advices. Thanks to Renaud Marty for his active involvement in testing AtmoSwing against his code base.

Thanks to the Swiss Federal Office for Environment (FOEV), the Roads and Water courses Service, Energy and Water Power Service of the Wallis Canton and the Water, Land and Sanitation Service of the Vaud Canton who financed the MINERVE project which started this research. NCEP reanalysis data provided by the NOAA/OAR/ESRL PSD, Boulder, Colorado, USA, from their Web site at http://www.esrl.noaa.gov/psd/. Precipitation time series provided by MeteoSwiss. 
\end{acknowledgements}

%% REFERENCES

%% The reference list is compiled as follows:

%\begin{thebibliography}{}
%
%\bibitem[AUTHOR(YEAR)]{LABEL}
%REFERENCE 1
%
%\bibitem[AUTHOR(YEAR)]{LABEL}
%REFERENCE 2
%
%\end{thebibliography}

\bibliographystyle{copernicus}
\bibliography{../_refs/_articles-atmoswing}

%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
%% authors experienced with BibTeX only have to include the following two lines:
%%
%% \bibliographystyle{copernicus}
%% \bibliography{example.bib}
%%
%% URLs and DOIs can be entered in your BibTeX file as:
%%
%% URL = {http://www.xyz.org/~jones/idx_g.htm}
%% DOI = {10.5194/xyz}

%% LITERATURE CITATIONS
%%
%% command                        & example result
%% \citet{jones90}|               & Jones et al. (1990)
%% \citep{jones90}|               & (Jones et al., 1990)
%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
%% \citeauthor{jones90}|          & Jones et al.
%% \citeyear{jones90}|            & 1990

%% FIGURES

%% ONE-COLUMN FIGURES

%%f
%\begin{figure}[t]
%\includegraphics[width=8.3cm]{FILE NAME}
%\caption{TEXT}
%\end{figure}
%
%%% TWO-COLUMN FIGURES
%
%%f
%\begin{figure*}[t]
%\includegraphics[width=12cm]{FILE NAME}
%\caption{TEXT}
%\end{figure*}
%
%
%%% TABLES
%%%
%%% The different columns must be seperated with a & command and should
%%% end with \\ to identify the column brake.
%
%%% ONE-COLUMN TABLE
%
%%t
%\begin{table}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{table}
%
%%% TWO-COLUMN TABLE
%
%%t
%\begin{table*}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{table*}
%
%
%%% NUMBERING OF FIGURES AND TABLES
%%%
%%% If figures and tables must be numbered 1a, 1b, etc. the following command
%%% should be inserted before the begin{} command.
%
%\addtocounter{figure}{-1}\renewcommand{\thefigure}{\arabic{figure}a}
%
%
%%% MATHEMATICAL EXPRESSIONS
%
%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
%%%
%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
%%% Vectors are identified in bold italic font using \vec{x}
%%% Matrices are identified in bold roman font
%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
%%% The character * should not be applied as mutliplication sign
%
%
%%% EQUATIONS
%
%%% Single-row equation
%
%\begin{equation}
%
%\end{equation}
%
%%% Multiline equation
%
%\begin{align}
%& 3 + 5 = 8\\
%& 3 + 5 = 8\\
%& 3 + 5 = 8
%\end{align}
%
%
%%% MATRICES
%
%\begin{matrix}
%x & y & z\\
%x & y & z\\
%x & y & z\\
%\end{matrix}
%
%
%%% ALGORITHM
%
%\begin{algorithm}
%\caption{…}
%\label{a1}
%\begin{algorithmic}
%…
%\end{algorithmic}
%\end{algorithm}
%
%
%%% CHEMICAL FORMULAS AND REACTIONS
%
%%% For formulas embedded in the text, please use \chem{}
%
%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
%
%\begin{reaction}
%%% \rightarrow should be used for normal (one-way) chemical reactions
%%% \rightleftharpoons should be used for equilibria
%%% \leftrightarrow should be used for resonance structures
%\end{reaction}
%
%
%%% PHYSICAL UNITS
%%%
%%% Please use \unit{} and apply the exponential notation

\end{document}